---
title: "Module_3_Assignment"
author: Jiayi (Joey) Liu, Luya Wang, Zexuan (Dante) Song
output:
  html_document:
    df_print: paged
---

## Load Libraries 

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(tidymodels)
library(janitor)
library(skimr)
library(kableExtra)
library(GGally)
library(vip)        
library(fastshap)   
library(MASS)
library(ISLR)
library(tree)
library(ggplot2)
library(dplyr)
library(lubridate)
library(imputeTS)
library(forecast)
library(urca)
library(astsa)
library(fpp2)
```

## Load Data
Read the .csv file into R.
```{r, message=FALSE,warning=FALSE}
data1 <- read_csv("arma_series01.csv") %>%
  clean_names()

data2 <- read_csv("arma_series02.csv") %>%
  clean_names()

data3 <- read_csv("arma_series03.csv") %>%
  clean_names()

data4 <- read_csv("arma_series04.csv") %>%
  clean_names()

data5 <- read_csv("arma_series05.csv") %>%
  clean_names()

data6 <- read_csv("arma_series06.csv") %>%
  clean_names()

data7 <- read_csv("arma_series07.csv") %>%
  clean_names()

```

```{r}
head(data1) 
head(data2) 
head(data3) 
head(data4) 
head(data5) 
head(data6) 
head(data7) 
```


```{r}
skim(data1)
skim(data2)
skim(data3)
skim(data4)
skim(data5)
skim(data6)
skim(data7)
# no missing values 
```

## clean data by deleting "date"


```{r}
data1 <- subset(data1, select = c(y))
data2 <- subset(data2, select = c(y))
data3 <- subset(data3, select = c(y))
data4 <- subset(data4, select = c(y))
data5 <- subset(data5, select = c(y))
data6 <- subset(data6, select = c(y))
data7 <- subset(data7, select = c(y))



```


## Create time series object and plot time series
```{r}

# Create time series object and plot time series
# data 1 - monthly index
ts1 <- ts(data1, start=c(1992,5), frequency = 12)
plot(ts1) 

# data 2 - daily index

as.numeric(as.Date("2008-06-15") - as.Date("2008-01-01"))
ts2 <- ts(data2, start=c(2008,166), frequency = 365)
plot(ts2) 


# data 3 - monthly index
ts3 <- ts(data3, start=c(1992,5), frequency = 12)
plot(ts3) 


# data 4 - daily index
as.numeric(as.Date("2005-05-08") - as.Date("2005-01-01"))
as.numeric(as.Date("2005-11-23") - as.Date("2005-01-01"))
ts4 <- ts(data4, start=c(2005,127), end = c(2005,326), frequency = 365)
plot(ts4) 

# data 5  - monthly index
ts5 <- ts(data5, start=c(1992,5), frequency = 12)
plot(ts5) 

# data 6  - monthly index
ts6 <- ts(data6, start=c(1992,5), frequency = 12)
plot(ts6)

# data 7  - monthly index
ts7 <- ts(data7, start=c(1993,1), frequency = 12)
plot(ts7) 


```


## Compute the Ljung Box Test for White Noise
```{r}


# test the residuals for white noise

Box.test(data1, lag=8, fitdf=0, type="Lj")
Box.test(data2, lag=8, fitdf=0, type="Lj")
Box.test(data3, lag=8, fitdf=0, type="Lj")
Box.test(data4, lag=8, fitdf=0, type="Lj")
Box.test(data5, lag=8, fitdf=0, type="Lj")
Box.test(data6, lag=8, fitdf=0, type="Lj")
Box.test(data7, lag=8, fitdf=0, type="Lj")

# P-value for all the 7 series < 0.05, reject the Null hypothesis, conclude that these series are NOT white noise 
```


# Plot ACF and PACF for data 1
# ADF test for stationarity
# Use the single mean Version of the Test


```{r, warning=FALSE, message=FALSE}

forecast::ggAcf(ts1)

forecast::ggPacf(ts1)


## drift - single mean, none - 0 mean, trend - trend
ts1_df <- ur.df(ts1, type = "drift")
summary(ts1_df)

#P-value of z.lag.1 is smaller than 0.05, reject H0, ts1 is stationary.
# d=0
```

## Estimate an AR(1) Model

```{r}
# Use sarima function which automatically computes p-values for parameter estimates and produces residual diagnostics

# PACF drops off quickly, start with p=1

fit1_AR1 <- sarima(ts1, 1, 0, 0)
summary(fit1_AR1)
fit1_AR1

# This is our final model (p=1, d=0, q=0)
# all terms are significant: p-value =0 < 0.05;
# Residuals white noice: ACF of residuals is within the boundary and P values of Ljung-Box are out of the line

```
```{r}
fitauto1<- auto.arima(ts1)
summary(fitauto1)
checkresiduals(fitauto1)

```


# Plot ACF and PACF for data 2
# ADF test for stationarity 2
# Use the single mean Version of the Test

```{r, warning=FALSE, message=FALSE}

forecast::ggAcf(ts2)

forecast::ggPacf(ts2)


## drift - single mean, none - 0 mean, trend - trend
ts2_df <- ur.df(ts2, type = "drift")
summary(ts2_df)

#P-value of z.lag.1 is smaller than 0.05, reject H0, ts2 is stationary.
# d=0
```

## Estimate an AR(2) Model

```{r}
# Use sarima function which automatically computes p-values for parameter estimates and produces residual diagnostics

# PACF quickly decays to 0, start with trying p =2 

fit2_AR2 <- sarima(ts2, 2, 0, 0)
summary(fit2_AR2)
fit2_AR2

## This is my final model for ts2 (p=2, d=0, q=0)
# All terms are significant: ar1, ar2 p-value =0<0.05
# Residuals are white noise 
```

#auto arima check
```{r}
fitauto2<- auto.arima(ts2)
summary(fitauto2)
checkresiduals(fitauto2)
```



# Plot ACF and PACF for data 3
# ADF test for stationarity 3
# Use the single mean Version of the Test

```{r, warning=FALSE, message=FALSE}

forecast::ggAcf(ts3)

forecast::ggPacf(ts3)


## drift - single mean, none - 0 mean, trend - trend
ts3_df <- ur.df(ts3, type = "drift")
summary(ts3_df)

#P-value of z.lag.1 is smaller than 0.05, reject H0, ts is stationarity.
# d=0

```

## Estimate an MA(2) Model

```{r}
# ACF drops off quickly, start with trying q=1
fit3_MA1 <- sarima(ts3, 0, 0, 1)
summary(fit3_MA1)
fit3_MA1

##This is our final model(p=0, d=0, q=1)
# All terms are significant 
# Residuals are white noise 

```

#auto arima check
```{r}
fitauto3<- auto.arima(ts3)
summary(fitauto3)
checkresiduals(fitauto3)
```



# Plot ACF and PACF for data 4
# ADF test for stationarity 4
# Use the single mean Version of the Test


```{r, warning=FALSE, message=FALSE}

forecast::ggAcf(ts4)

forecast::ggPacf(ts4)


## drift - single mean, none - 0 mean, trend - trend
ts4_df <- ur.df(ts4, type = "drift")
summary(ts4_df)

#P-value of z.lag.1 is smaller than 0.05, reject H0, ts is stationarity.
# d=0

```

## Estimate an MA(2) Model

```{r}
# ACF drops off quickly, start with trying q=2
fit4_MA2 <- sarima(ts4, 0, 0, 2)
summary(fit4_MA2)
fit4_MA2

# This is our final model (p=0,d=0, q=2)
# All terms are significant 
# Residual are white noise 

```


#auto arima check
```{r}
fitauto4<- auto.arima(ts4)
summary(fitauto4)
checkresiduals(fitauto4)
```



# Plot ACF and PACF for data 5
# ADF test for stationarity 5
# Use the single mean Version of the Test


```{r, warning=FALSE, message=FALSE}

forecast::ggAcf(ts5)

forecast::ggPacf(ts5)



## drift - single mean, none - 0 mean, trend - trend
ts5_df <- ur.df(ts5, type = "drift")
summary(ts5_df)

#P-value of z.lag.1 is smaller than 0.05, reject H0, ts is stationarity.
# d=0

```

## Estimate an AR(5) Model

```{r}
# lag 1 PACF is positive, start with trying p=1
# The auto.arima gives me (1,0,2), then I tried (1,0,1). The result is still not good. so I chose the number gave by auto.arima.

fit5_AR1 <- sarima(ts5, 1, 0, 2)
summary(fit5_AR1)
fit5_AR1

# This is our final model (p=1, d=0, q=2)
# All terms are significant 
# Residuals are white noise 

```


#auto arima check
```{r}
fitauto5<- auto.arima(ts5)
summary(fitauto5)
checkresiduals(fitauto5)
```



# Plot ACF and PACF for data 6
# ADF test for stationarity 6
# Use the single mean Version of the Test


```{r, warning=FALSE, message=FALSE}

forecast::ggAcf(ts6)

forecast::ggPacf(ts6)



## drift - single mean, none - 0 mean, trend - trend
ts6_df <- ur.df(ts6, type = "drift")
summary(ts6_df)

#P-value of z.lag.1 is smaller than 0.05, reject H0, ts is stationarity.
# d=0
```

## Estimate an MA(2) Model

```{r}
# lag1 PACF is negative, start with q=1
# Then I tried q=2. The result is not ideal. Then I tries (1,0,2), still not ideal.
# Then auto.arima gave me (2,0,1)

fit6_MA2 <- sarima(ts6, 2,0,1)
summary(fit6_MA2)
fit6_MA2

# This is our final model, using the number generated by AUTO.arima(p=2, d=0, q=1)
# All terms are significant 
# Residuals are significant

```

#auto arima check
```{r}
fitauto6<- auto.arima(ts6)
summary(fitauto6)
checkresiduals(fitauto6)
```


# Plot ACF and PACF for data 7
# ADF test for stationarity 7
# Use the single mean Version of the Test


```{r, warning=FALSE, message=FALSE}

forecast::ggAcf(ts7)

forecast::ggPacf(ts7)


# ADF test for stationarity 7
# Use the single mean Version of the Test

## drift - single mean, none - 0 mean, trend - trend
ts7_df <- ur.df(ts7, type = "drift")
summary(ts7_df)



```

## Estimate an AR(7) Model

```{r}
# PACF quickly drops off, try with p=1
# Then I tried adding q=1

fit7_AR1 <- sarima(ts7, 1, 0, 1)
summary(fit7_AR1)
fit7_AR1

## This is our final model(p=1,d=0,q=1)
# All terms are significant 
# Residuals are white noise 

```

#auto arima check
```{r}
fitauto7<- auto.arima(ts7)
summary(fitauto7)
checkresiduals(fitauto7)
```

